\chapter{Technologies}
\subsection{CUDA}

CUDA (Compute Unified Device Architecture) is a parallel computing platform and application programming interface (API) developed by NVIDIA. It enables developers to utilize the power of NVIDIA GPUs for general-purpose computing tasks. By leveraging CUDA, applications can perform computations faster by distributing tasks across thousands of GPU cores. It is widely used in fields such as scientific computing, machine learning, and real-time rendering.


\subsubsection{NVIDIA Nsight Compute}
NVIDIA Nsight Compute is a performance analysis tool designed for CUDA applications. It provides detailed metrics and insights into GPU kernel execution, enabling developers to identify bottlenecks and optimize their code. This tool is essential for understanding and improving the performance of GPU-accelerated applications.

\subsubsection{NVIDIA Nsight Systems}
NVIDIA Nsight Systems is a system-wide performance analysis tool that helps developers identify performance issues across the CPU and GPU. It provides a holistic view of application behavior, enabling optimization at both the system and application levels.

In the results section, we will analyze our findings using these tools to evaluate the performance of our CUDA-based implementations.

\subsection{OpenCV}

OpenCV (Open Source Computer Vision Library) is an open-source library designed for real-time computer vision and image processing tasks. It provides a wide range of tools and functions for image manipulation, feature extraction, object detection, and more. OpenCV is widely adopted in industries such as robotics, medical imaging, and autonomous vehicles due to its versatility and robust performance.

Some key features of OpenCV include:
\begin{itemize}
\item \textbf{Wide Functionality:} OpenCV supports tasks such as image filtering, geometric transformations, and object recognition.
\item \textbf{Cross-Platform Support:} OpenCV works on multiple platforms, including Windows, Linux, macOS, and Android.
\item \textbf{Hardware Acceleration:} OpenCV integrates with CUDA and OpenCL to accelerate processing on compatible hardware.
\end{itemize}

OpenCV is written in C++ but provides bindings for Python, Java, and other languages, making it accessible to a wide range of developers. With its comprehensive functionality and active community, OpenCV remains a cornerstone of computer vision development.

